---
title: "TidyText"
# subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
subtitle: "An introduction to text analysis in R"
author: "Nathan Bracken"
date: "06/02/2021"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true 
      highlightSpans: true
      countIncrementalSlides: false
      keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages, cache=FALSE, message=FALSE}
pacman::p_load(tidyverse, DBI, bigrquery,dbplyr, tokenizers, tidytext,stopwords)
```

# Text Analysis/Sentiment Anlysis/NLP?

- Allowing Computers to work with human language 
- String down to Binary!
- We can identify implied meaning behind words
- Useful for aiding in the understanding of perception
- Different lexicons and different ways to evaluate the language
- Different analysis levels

---

#Today we are focusing on TidyText

- There are many other packages useful for text analysis 
- Allows for us to text mine and use Tidyverse wrangling tools
- Great because we are very familiar with the Dplyr wrangling methods

---

# Using BigQuery to Acquire Data

```{r echo = TRUE, message=FALSE}
# library(DBI, bigrquery, dbplyr)

billing_id = Sys.getenv("GCE_TIDYTEXT_PROJECT_ID")

books_bq_con = dbConnect(bigrquery::bigquery(),
                project ="gdelt-bq", 
                dataset = "internetarchivebooks",
                billing = billing_id,
                use_legacy_sql = FALSE)

books_table_1920 = tbl(books_bq_con,'1920')
```

---
#So Many Tables!

```{r}
dbListTables(books_bq_con)
```

---
#Filtering

```{r echo = TRUE}
#library(tidyverse)
text_query = books_table_1920 %>% 
  filter(BookMeta_Language == "eng") %>% 
  filter(grepl(BookMeta_Subjects,'poetry')) %>%
  collect()
```

```{r}
dbDisconnect(books_bq_con)
```

---

#Creating a table

```{r}
tidy_poetry = text_query %>% 
              select(BookMeta_Title, BookMeta_Subjects, BookMeta_FullText) 

tidy_poetry
```

---

#Tokenizing?

```{r echo = TRUE}
#library(tokenizers)
tokenized_poetry = tidy_poetry %>% 
    group_by(BookMeta_Title) %>%
    unnest_tokens(word, BookMeta_FullText)
```

---

#Tokenized Meta Text 

```{r}
tokenized_poetry
```

---

#Counting Words

```{r}
##general sentiment of all poetry
##general count of the words
counted_poetry = tokenized_poetry %>%
  ungroup() %>% 
  count(word, sort = TRUE)
counted_poetry
```

---

#Count of Words by Book

```{r}
##count of words by book
counted_poetrybook = tokenized_poetry %>% 
  count(word, sort = TRUE)
counted_poetrybook
```

---

#Stopping Words

```{r echo = TRUE}
##library(stopwords)
getstop_poetry = tokenized_poetry %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE)
```
- Get_stopwords stops the reading of certain words
- There are many different stopword methods (Snowball, SMART, marimo, etc.) 
---

#This looks a bit better

```{r}
getstop_poetry
```

---

#Calculating sentiment

```{r echo = TRUE}
##library(tidyr)
##1920 poetry sentiment 
sentiment_1920 = getstop_poetry %>% 
    inner_join(get_sentiments("bing"), by = "word") %>% 
  count(BookMeta_Title, sentiment) %>% 
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

---

#Sentiment Results Positive

```{r}
sentiment_1920 %>% arrange(desc(sentiment))
```


---

#Sentiment Results Negative

```{r}
sentiment_1920 %>% arrange(sentiment)
```

---

#Plot of Sentiment Levels

```{r}
#plot need to add labels and fix the scaling
sentiment_plot = sentiment_1920 %>% 
  ggplot(aes(BookMeta_Title, sentiment)) +
        geom_bar(stat = "identity", show.legend = FALSE) + 
        geom_text(aes(label = sentiment), vjust = 1.5) +
        theme(axis.text.x=element_blank())

sentiment_plot
```

---

#Resources

##Here are some useful resources if you want to get started on NLP!

- Julia Silge's Github / Implementation on books by Jane Austen
  - https://github.com/juliasilge/tidytext/
  - https://juliasilge.github.io/tidytext/
- Tidy Text Mining with R book
  - https://www.tidytextmining.com/
- String Processing Package: stringi
  - https://stringi.gagolewski.com/
- Tokenizers Package
  - https://docs.ropensci.org/tokenizers/
- CRAN Task View NLP
  - https://cran.r-project.org/web/views/NaturalLanguageProcessing.html

